---
#PROLOG   ###############################################################

#PROJECT: Final Project
#PURPOSE: Cox PH Regression
#DIR:     /Users/Arash/CancerAmenability1
#DATA:    BRCA_SEER_SURV
#AUTHOR:  Arash Delavar
#CREATED: April 8th, 2018
#LATEST:  May 9th, 2018
#NOTES:   This code was written to perform the final project, "Racial Childhood Cancer Survival Disparities by Cancer Amenability"

#PROLOG   ###############################################################
---
Load Packages
```{r setup, include=FALSE}
#install.packages("survminer")#for pairwise diffs
#install.packages("lmtest")
#install.packages("stargazer")
library(readr) #for read txt file
library(survival) #for calculating KM values
library(survminer)#for pairwise diffs
library(ggfortify) #for KM curves
library(readxl) # for reading in excel file
library(ggplot2) # for plotting KM curve
library(tidyverse) # for various packages
library(survminer)#for pairwise diffs
library(lmtest) #model comparison
library(stargazer) #models presentation
```

Load data and merge
```{r}
#load data
CancerAmenability <- read.csv ("/Users/Arash/downloads/CancerAmenability2.csv")
RSR <- read.csv("/Users/Arash/downloads/RSR11.csv")
m <- merge(CancerAmenability,RSR)
```

Clean data
```{r}
#new column names
colnames(m)[colnames(m)=="Median.family.income..in.tens..2000"]<-"income"
colnames(m)[colnames(m)=="X....High.school.education.2000"]<-"education"
colnames(m)[colnames(m)=="Age.at.diagnosis"]<- "ageatd"
colnames(m)[colnames(m)=="Race.and.origin.recode..NHW..NHB..NHAIAN..NHAPI..Hispanic."]<-"Race"
colnames(m)[colnames(m)=="SEER.cause.specific.death.classification"]<-"death"
colnames(m)[colnames(m)=="Insurance.Recode..2007.."]<-"insurance"
#rid missing
m <- m[which(m$Race<9),]
m <- m[which(m$death<8),]
m <- m[which(m$`Survival.months`<9999),]
m <- m[which(m$Relative>0),]
m[which(m$`education`==16382),]<- NA
m <- m[which(m$income!=16382),]

#divide into 5 household income categories: from low to high
c <- quantile(m$income,probs = c(.2,.4,.6,.8,1),na.rm = T)
c
#divide education into 5 categories: from low to high
e <- quantile(m$education,probs = c(.2,.4,.6,.8,1),na.rm = T)
e
```

Arash recodes for m
```{r}
#recode race
m$Race <- factor(m$Race,levels = c(1:5),labels = c("NHW","NHB","NHAIAN","NHAPI","Hispanic"))

#recode sex
m$Sex <- factor(m$Sex,levels = c(1,2),labels = c("Male","Female"))

#recode age
m$Age[m$`ageatd`<5]<-0
m$Age[m$`ageatd`<10 & m$`ageatd`>=5]<-1
m$Age[m$`ageatd`<15 & m$`ageatd`>=10]<-2
m$Age[m$`ageatd`<20 & m$`ageatd`>=15]<-3
m$Age <- factor(m$Age,levels=c(0:3),labels=c("0-4","5-9","10-14","15-19"))

#recode income
m$income_Cat[m$income<c[1]]<-1
m$income_Cat[m$income>=c[1] & m$income<c[2]]<-2
m$income_Cat[m$income>=c[2] & m$income<c[3]]<-3
m$income_Cat[m$income>=c[3] & m$income<c[4]]<-4
m$income_Cat[m$income>=c[4] ] <- 5

#recode education
m$edu_Cat[m$education<e[1]]<-1
m$edu_Cat[m$education>=e[1] & m$education<e[2]]<-2
m$edu_Cat[m$education>=e[2] & m$education<e[3]]<-3
m$edu_Cat[m$education>=e[3] & m$education<e[4]]<-4
m$edu_Cat[m$education>=e[4] ] <- 5

#divide into 3 amenability categories (m1-m3) and another 3 for 2007+ (m4-m6)
m1 <- m[which(m$Relative<.7),] #non-amenable
m2 <- m[which(m$Relative>=0.7 & m$Relative<=0.85),] #part-amenable
m3 <- m[which(m$Relative>=.85),] #amenable
m4 <- m[which(m$Relative<.7 & m$insurance<14),] #non-amenable
m5 <- m[which(m$Relative>=0.7 & m$Relative<=0.85 & m$insurance<14),] #part-amenable
m6 <- m[which(m$Relative>=.85 & m$insurance<14),] #amenable



#Cancer type breakdown

#I Leukemias, myeloproliferative diseases, and myelodysplastic diseases
m$Cancer_Type[m$ICCCName=="I(a.1) Precursor cell leukemias"|
                m$ICCCName=="I(a.2) Mature B-cell leukemias"|
                m$ICCCName=="I(a.3) Mature T-cell and NK cell leukemias"|
                m$ICCCName=="I(a.4) Lymphoid leukemia, NOS"|
                m$ICCCName=="I(b) Acute myeloid leukemias"|
                m$ICCCName=="I(c) Chronic myeloproliferative diseases"|
                m$ICCCName=="I(d) Myelodysplastic syndrome and other myeloproliferative"|
                m$ICCCName=="I(e) Unspecified and other specified leukemias"]<-1

#II Lymphomas and reticuloendothelial neoplasms
m$Cancer_Type[m$ICCCName=="II(a) Hodgkin lymphomas"|
                m$ICCCName=="II(b.1) Precursor cell lymphomas"|
                m$ICCCName=="II(b.2) Mature B-cell lymphomas except Burkitt lymphoma"|
                m$ICCCName=="II(b.3) Mature T-cell and NK-cell lymphomas"|
                m$ICCCName=="II(b.4) Non-Hodgkin lymphomas, NOS"|
                m$ICCCName=="II(c) Burkitt lymphoma"|
                m$ICCCName=="II(d) Miscellaneous lymphoreticular neoplasms"|
                m$ICCCName=="II(e) Unspecified lymphomas"]<-2

# III CNS and miscellaneous intracranial and intraspinal neoplasms
m$Cancer_Type[m$ICCCName=="III(a.1) Ependymomas"|
                m$ICCCName=="III(a.2) Choroid plexus tumor"|
                m$ICCCName=="III(b) Astrocytomas"|
                m$ICCCName=="III(c.1) Medulloblastomas"|
                m$ICCCName=="III(c.2) PNET"|
                m$ICCCName=="III(c.3) Medulloepithelioma"|
                m$ICCCName=="III(c.4) Atypical teratoid/rhabdoid tumor"|
                m$ICCCName=="III(d.1) Oligodendrogliomas"|
                m$ICCCName=="III(d.2) Mixed and unspecified gliomas"|
                m$ICCCName=="III(d.3) Neuroepithelial glial tumors of uncertain orig"|
                m$ICCCName=="III(e.1) Pituitary adenomas and carcinomas"|
                m$ICCCName=="III(e.2) Tumors of sellar region (craniopharyngiomas)"|
                m$ICCCName=="III(e.3) Pineal parenchymal tumors"|
                m$ICCCName=="III(e.4) Neuronal and mixed neuronal-glial tumors"|
                m$ICCCName=="III(e.5) Meningiomas"|
                m$ICCCName=="III(f) Unspecified intracranial and intraspinal neoplasms"]<-3

# IV Neuroblastoma and other peripheral nervous cell tumors
m$Cancer_Type[m$ICCCName=="IV(a) Neuroblastoma and ganglioneuroblastoma"|
                m$ICCCName=="IV(b) Other peripheral nervous cell tumors"]<-4

# V Retinoblastoma
m$Cancer_Type[m$ICCCName=="V Retinoblastoma"]<-5

# VI Renal tumors
m$Cancer_Type[m$ICCCName=="VI(a.1) Nephroblastoma"|
                m$ICCCName=="VI(a.2) Rhabdoid renal tumor"|
                m$ICCCName=="VI(a.3) Kidney sarcomas"|
                m$ICCCName=="VI(a.4) pPNET of kidney"|
                m$ICCCName=="VI(b) Renal carcinomas"|
                m$ICCCName=="VI(c) Unspecified malignant renal tumors"]<-6

# VII Hepatic tumors
m$Cancer_Type[m$ICCCName=="VII(a) Hepatoblastoma"|
                m$ICCCName=="VII(b) Hepatic carcinomas"|
                m$ICCCName=="VII(c) Unspecified malignant hepatic tumors"]<-7

#VIII Malignant bone tumors
m$Cancer_Type[m$ICCCName=="VIII(a) Osteosarcomas"|
                m$ICCCName=="VIII(b) Chondrosarcomas"|
                m$ICCCName=="VIII(c.1) Ewing tumor and Askin tumor of bone"|
                m$ICCCName=="VIII(c.2) pPNET of bone"|
                m$ICCCName=="VIII(d.1) Malignant fibrous neoplasms of bone"|
                m$ICCCName=="VIII(d.2) Malignant chordomas"|
                m$ICCCName=="VIII(d.3) Odontogenic malignant tumors"|
                m$ICCCName=="VIII(d.4) Miscellaneous malignant bone tumors"|
                m$ICCCName=="VIII(e) Unspecified malignant bone tumors"]<-8

#IX Soft tissue and other extraosseous sarcomas
m$Cancer_Type[m$ICCCName=="IX(a) Rhabdomyosarcomas"|
                m$ICCCName=="IX(b.1) Fibroblastic and myofibroblastic tumors"|
                m$ICCCName=="IX(b.2) Nerve sheath tumors"|
                m$ICCCName=="IX(b.3) Other fibromatous neoplasms"|
                m$ICCCName=="IX(c) Kaposi sarcoma"|
                m$ICCCName=="IX(d.1) Ewing tumor and Askin tumor of soft tissue"|
                m$ICCCName=="IX(d.2) pPNET of soft tissue"|
                m$ICCCName=="IX(d.3) Extrarenal rhabdoid tumor"|
                m$ICCCName=="IX(d.4) Liposarcomas"|
                m$ICCCName=="IX(d.5) Fibrohistiocytic tumors"|
                m$ICCCName=="IX(d.6) Leiomyosarcomas"|
                m$ICCCName=="IX(d.7) Synovial sarcomas"|
                m$ICCCName=="IX(d.8) Blood vessel tumors"|
                m$ICCCName=="IX(d.9) Osseous & chondromatous neoplasms of soft tissue"|
                m$ICCCName=="IX(d.10) Alveolar soft parts sarcoma"|
                m$ICCCName=="IX(d.11) Miscellaneous soft tissue sarcomas"|
                m$ICCCName=="IX(e) Unspecified soft tissue sarcomas"]<-9

#X Germ cell tumors, trophoblastic tumors, and neoplasms of gonads
m$Cancer_Type[m$ICCCName=="X(a.1) Intracranial & intraspinal germinomas"|
m$ICCCName=="X(a.2) Intracranial & intraspinal teratomas"|
m$ICCCName=="X(a.3) Intracranial & intraspinal embryonal carcinomas"|
m$ICCCName=="X(a.4) Intracranial & intraspinal yolk sac tumor"|
m$ICCCName=="X(a.5) Intracranial & intraspinal choriocarcinoma"|
m$ICCCName=="X(a.6) Intracranial & intraspinal tumors of mixed forms"|
m$ICCCName=="X(b.1) Germinomas: extracranial/extragonadal"|
m$ICCCName=="X(b.2) Malignant teratomas: extracranial/extragonadal"|
m$ICCCName=="X(b.3) Embryonal carcinomas: extracranial/extragonadal"|
m$ICCCName=="X(b.4) Yolk sac tumor: extracranial/extragonadal"|
m$ICCCName=="X(b.5) Choriocarcinomas: extracranial/extragonadal"|
m$ICCCName=="X(b.6) Other mixed germ cell: extracranial/extragonadal"|
m$ICCCName=="X(c.1) Malignant gonadal germinomas"|
m$ICCCName=="X(c.2) Malignant gonadal teratomas"|
m$ICCCName=="X(c.3) Gonadal embryonal carcinomas"|
m$ICCCName=="X(c.4) Gonadal yolk sac tumor"|
m$ICCCName=="X(c.5) Gonadal choriocarcinoma"|
m$ICCCName=="X(c.6) Malignant gonadal tumors of mixed forms"|
m$ICCCName=="X(d) Gonadal carcinomas"|
m$ICCCName=="X(e) Other and unspecified malignant gonadal tumors"]<-10

#VI Other malignant epithelial neoplasms and malignant melanomas
m$Cancer_Type[m$ICCCName=="XI(a) Adrenocortical carcinomas"|
m$ICCCName=="XI(b) Thyroid carcinomas"|
m$ICCCName=="XI(c) Nasopharyngeal carcinomas"|
m$ICCCName=="XI(d) Malignant melanomas"|
m$ICCCName=="XI(e) Skin carcinomas"|
m$ICCCName=="XI(f.1) Carcinomas of salivary glands"|
m$ICCCName=="XI(f.2) Carcinomas of colon and rectum"|
m$ICCCName=="XI(f.3) Carcinomas of appendix"|
m$ICCCName=="XI(f.4) Carcinomas of lung"|
m$ICCCName=="XI(f.5) Carcinomas of thymus"|
m$ICCCName=="XI(f.6) Carcinomas of breast"|
m$ICCCName=="XI(f.7) Carcinomas of cervix uteri"|
m$ICCCName=="XI(f.8) Carcinomas of bladder"|
m$ICCCName=="XI(f.9) Carcinomas of eye"|
m$ICCCName=="XI(f.10) Carcinomas of other specified sites"|
m$ICCCName=="XI(f.11) Carcinomas of unspecified site"]<-11

#VII Other and unspecified malignant neoplasms
m$Cancer_Type[m$ICCCName=="XII(a.1) Gastrointestinal stromal tumor"|
m$ICCCName=="XII(a.2) Pancreatoblastoma"|
m$ICCCName=="XII(a.3) Pulmonary blastoma and pleuropulmonary blastoma"|
m$ICCCName=="XII(a.4) Other complex mixed and stromal neoplasms"|
m$ICCCName=="XII(a.5) Mesothelioma"|
m$ICCCName=="XII(a.6) Other specified malignant tumors"|
m$ICCCName=="XII(b) Other unspecified malignant tumors"]<-12

#Not classified by ICCC or in situ
m$Cancer_Type[m$ICCCName=="Not classified by ICCC or in situ"]<-13

m$Cancer_Type <- factor(m$Cancer_Type,levels = c(1:13),
                        labels = c("Leukemias, myeloproliferative diseases, and myelodysplastic diseases",
                                   "Lymphomas and reticuloendothelial neoplasms",
                                   "CNS and miscellaneous intracranial and intraspinal neoplasms",
                                   "Neuroblastoma and other peripheral nervous cell tumors",
                                   "Retinoblastoma",
                                   "Renal tumors",
                                   "Hepatic tumors",
                                   "Malignant bone tumors",
                                   "Soft tissue and other extraosseous sarcomas",
                                   "Germ cell tumors, trophoblastic tumors, and neoplasms of gonads",
                                   "Other malignant epithelial neoplasms and malignant melanomas",
                                   "Other and unspecified malignant neoplasms",
                                   "Not classified by ICCC or in situ"))

#unique(m$ICCCName)
#sum(m$ICCCName== "NA")
```
```

Table 1
```{r}
#Descriptive demographics
table(m$Race,m$death)
table(m$Sex,m$death)
table(m$Age,m$death)
table(m$income_Cat,m$death)
table(m$Cancer_Type,m$death)
#total by race
table(m$Race)
table(m1$Race)
table(m2$Race)
table(m3$Race)
table(m4$Race)
table(m5$Race)
table(m6$Race)
#death counts by race
table(m[which(m$`death`==1),]$Race)
table(m1[which(m1$`death`==1),]$Race)
table(m2[which(m2$`death`==1),]$Race)
table(m3[which(m3$`death`==1),]$Race)
table(m4[which(m4$`death`==1),]$Race)
table(m5[which(m5$`death`==1),]$Race)
table(m6[which(m6$`death`==1),]$Race)
```

Table 2
```{r}
#Non-Amenable
category.mod1<-coxph(Surv(m1$`Survival.months`,m1$`death`)~Race+edu_Cat+income_Cat+Age+Sex, m1, ties="breslow")

#Part-Amenable
category.mod2<-coxph(Surv(m2$`Survival.months`,m2$`death`)~Race+edu_Cat+income_Cat+Age+Sex, m2, ties="breslow")

#Amenable
category.mod3<-coxph(Surv(m3$`Survival.months`,m3$`death`)~Race+edu_Cat+income_Cat+Age+Sex, m3, ties="breslow")

summary(category.mod1)
summary(category.mod2)
summary(category.mod3)

#Non-Amenable +insurance
category.mod4<-coxph(Surv(m4$`Survival.months`,m4$`death`)~Race+edu_Cat+income_Cat+Age+Sex+insurance, m4, ties="breslow")

#Part-Amenable +insurance
category.mod5<-coxph(Surv(m5$`Survival.months`,m5$`death`)~Race+edu_Cat+income_Cat+Age+Sex+insurance, m5, ties="breslow")

#Amenable +insurance
category.mod6<-coxph(Surv(m6$`Survival.months`,m6$`death`)~Race+edu_Cat+income_Cat+Age+Sex+insurance, m6, ties="breslow")

summary(category.mod4)
summary(category.mod5)
summary(category.mod6)
```